---
title: "CDC Leading Causes of Death 1999-2016"
author: "Connor Sakmar"
date: "December 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(kableExtra)
library(tidytext)
library(knitr)
library(janitor)
library(lubridate)
library(stringr)
library(fs)
library(formattable)
library(ggplot2)
library(gganimate)
library(leaflet)
library(rgdal)
library(maps)
library(mapproj)
library(fiftystater)
library(sf)
library(tidyverse)
```

```{r, echo = FALSE}





leading_deaths <- read.csv("NCHS_-_Leading_Causes_of_Death__United_States.csv")

  




#This is the code for my Bar Graph, here we want the user to be able to compare all of the 
#causes next to each other.When put into shiny, In the shiny app, it is possible to choose 
#the year so that the causes can be compared for every year.

    myPlot1 <- leading_deaths %>%
      select(Year, Deaths, State, Cause.Name) %>%
    # This filter below will be added into shiny to make the year an important function in the shiny app:
    #  filter(Year == input$x) %>%
      
      #We want to compare just the ten causes and just the United States totals.
      #The US totals are more impactful, and All Causes data does not just ruin
      #scaling on the bar graph but the All Causes data is incomplete for the year 2016
      
      filter(State == "United States", Cause.Name != "All causes") %>%
      ggplot(aes(x = Cause.Name, y = Deaths, fill = Cause.Name)) + geom_bar(stat = "identity") +
      xlab("Year") + ylab("Number of Deaths") + 
      ggtitle("Leading Causes of Death in the United States 1999-2016") +
      
      #Getting rid of the bottom tick marks and cause names as they made the bottom 
      #of the graph cluttered and were also redundant. 
      theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
    
    #The deaths total is easier to read in thousands than in scientific notation
    #so this code converts the deaths into thousands below. 
    require(scales)
    myPlot1 + scale_y_continuous(labels = comma)
    
    
    
#This is the code for the Scatter Plot. Unlike the Bar Graph, we want the user to be able to 
#see the trends of each individual cause throughout the given year period 1999-2016. In
#the shiny app, this is made possible by a pull down menu to input each different cause of death. 

    
       myPlot2 <- leading_deaths %>%
         
      #Again, looking at just the United States totals.
      filter(State == "United States") %>%
         
      #This is the filter that put into the shiny app that allows
      #the user to input each different cause of death:
      #filter(Cause.Name == input$Cause.Name) %>%
         
      
      #If you run just this ggplot, it will give a strange scatterplot 
      #that includes all of the deaths and year. Once the filter is added
      #in shiny though, this scatterplot will make sense. 
      ggplot(aes(x = Year, y = Deaths)) + geom_point() + geom_smooth() 
       
#The deaths total is easier to read in thousands than in scientific notation
#so this code converts the deaths into thousands below. 
    
    require(scales)
    myPlot2 + scale_y_continuous(labels = comma)




  
```
```
```{r}

#This is the code tha used for the map. Most of this is put into a RDS.


#Rereading the leading deaths as it may be manipulated later in the code.
leading_deaths <- read.csv("NCHS_-_Leading_Causes_of_Death__United_States.csv")

#As mentioned earlier in the code, the 'All Causes' data for 2016 is incomplete
#in the CDC dataset, so it is easier to filter it out of leading_deaths at the very beginning of the 
#map code to avoid any future complications. 
leading_deaths <- leading_deaths %>%
filter(Cause.Name != 'All causes')

#The fiftystater library is loaded as it allows us to include all 50 states into the map.
#The library allows be to call the data with fifty_states, though it has to be converted it 
#to a sf data frame and saving it as sf_fifty.
(sf_fifty <- st_as_sf(fifty_states, coords = c("long", "lat")) %>% 
    
#Converting the sets of points within the data into polygons
    group_by(id, piece) %>% 
    summarize(do_union = FALSE) %>%
    st_cast("POLYGON") %>%
    
#Converting the polygons to multipolygons for states with discontinuous regions
    group_by(id) %>% 
    summarize())
   
#This retrieves a coordinate reference system. 
st_crs(sf_fifty) <- 4326

#Setting the states in leading death to lowercase so that the data can be joined with 
#sf_fifty properly
leading_deaths$id <- tolower(leading_deaths$State)
us_states_leading_deaths <- left_join(sf_fifty, leading_deaths)

#Creating a RDS file to put into shiny so that the code does not have to re-run in shiny
#every time the app is run. 
write_rds(us_states_leading_deaths, "us_states_leading_deaths.rds", compress = "none")

#This is the basic ggplot structure for the map. 
ggplot(data = us_states_leading_deaths) +
  geom_sf(aes(fill = Deaths))
  
```
  
  
